### 2020-10-27 latex/acrobot.tex
Started writing the acrobot constraint section
Need to describe the constrained dynamics in (qu,pu) and explain why we
can't use any form of Eavg before moving onto the proof.

While this is the most general representation of an acrobot, the dynamics
become unwieldy.
To make rigorous analysis of these dynamics more tractable, we begin by assuming
the acrobot is comprised of two massless rods of equal length \(l\), with equal
point masses \(m\) at the tips.
We call this a \textit{simple} acrobot, which is displayed in Figure
\ref{fig:simple-acrobot-model}.
We will also ignore any frictional forces at both the hip and shoulder joints. 
Finally, it is important to note that a real gymnast cannot swing their legs in
full circles; 
for this reason, we assume that \(q_a \in [-\bar{q}_a, \bar{q}_a]\) where 
\(\bar{q}_a \in ]-\pi, \pi[\).

@@

Let us turn next to the thesis of \citet{xingbo_thesis}, who designs a VHC to enforce a
so-called ``tap" motion with the purpose of injecting energy into the acrobot. 
First, he defines a compensator variable \(s\) which tracks \(\dot{q}_u\), so
that he can use the theory of VHCs with the extended configuration 
\((q_u,q_a,s)\).
He then finds \(h_1, h_2 \in \R_{>0}\) to define the
normalized radius \(\rho\) and normalized angle \(\xi\) in the
\((q_u, s)\)-plane.
These normalized variables are given by
\begin{align*}
    \rho &:= \sqrt{h_1 q_u^2 + h_2 s^2}
    , \\
    \xi &:= \arctan_2(h_2 s, h_1 q_u)
    . 
\end{align*}
He then sets the VHC to be \(h(q) = q_a - f_\text{rad}(\rho)f_\text{ang}(\xi)\)
with the control parameters \(\bar{q}_u\) and \(\rho_0\), where
\begin{align*}
    f_\text{rad}(\rho) &:= \tanh^2(\rho/\rho_0)
    , \\
    f_\text{ang}(\xi) &:= 
    \begin{cases}
        0 & -\pi < \xi \leq 0 \\
        \bar{q}_u \exp\left(1 - \frac{1}{1-(\frac{4\xi}{\pi} - 1)^2}\right) 
          & 0 < \xi \leq \frac{\pi}{2} \\
        0 & \frac{\pi}{2} < \xi \leq \pi
        .
    \end{cases}
\end{align*}

@@

\section{The Acrobot Constraint}
% TODO: Give the constraint, state the theorem, and show the outline of the
% proof (both as a sketch and in a figure)
\section{Proving the Acrobot Gains Energy}
% TODO: Prove this in sections

>>>>

While this is the most general representation of an acrobot, the dynamics
become unwieldy.
To make rigorous analysis of these dynamics more tractable, we begin by assuming
the acrobot is comprised of two massless rods of equal length \(l\), with equal
point masses \(m\) at the tips.
We call this a \textit{simple} acrobot, which is displayed in Figure
\ref{fig:simple-acrobot-model}.
We will also ignore any frictional forces at both the hip and shoulder joints. 
Finally, it is important to note that a real gymnast cannot swing their legs in
full circles; 
for this reason, we assume that \(q_a \in [-Q_a, Q_a]\) where 
\(Q_a \in ]-\pi, \pi[\).

@@

Let us turn next to the thesis of \citet{xingbo_thesis}, who designs a VHC to enforce a
so-called ``tap" motion with the purpose of injecting energy into the acrobot. 
First, he defines a compensator variable \(s\) which tracks \(\dot{q}_u\), so
that he can use the theory of VHCs with the extended configuration 
\((q_u,q_a,s)\).
He then finds \(h_1, h_2 \in \R_{>0}\) to define the
normalized radius \(\rho\) and normalized angle \(\xi\) in the
\((q_u, s)\)-plane.
These normalized variables are given by
\begin{align*}
    \rho &:= \sqrt{h_1 q_u^2 + h_2 s^2}
    , \\
    \xi &:= \arctan_2(h_2 s, h_1 q_u)
    . 
\end{align*}
He then sets the VHC to be \(h(q) = q_a - f_\text{rad}(\rho)f_\text{ang}(\xi)\)
with the control parameters \(\bar{q}_u\) and \(\rho_0\), where
\begin{align}
    \label{eqn:xingbo-frad}
    f_\text{rad}(\rho) &:= \tanh^2(\rho/\rho_0)
    , \\
    \label{eqn:xingbo-fang}
    f_\text{ang}(\xi) &:= 
    \begin{cases}
        0 & -\pi < \xi \leq 0 \\
        \bar{q}_u \exp\left(1 - \frac{1}{1-(\frac{4\xi}{\pi} - 1)^2}\right) 
          & 0 < \xi \leq \frac{\pi}{2} \\
        0 & \frac{\pi}{2} < \xi \leq \pi
        .
    \end{cases}
\end{align}

@@

\section{The Acrobot Constraint}
% TODO: Give the constraint, state the theorem, and show the outline of the
% proof (both as a sketch and in a figure)
One may be tempted to design a constraint \(q_a = f(\theta)\) 
with \(\theta := \arctan_2(p_u,q_u)\), since it was so effective for the VLP in
Chapter \ref{ch:vlp}.
It can be challenging to show this type of constraint is regular.
Recall that all constraints of the form \(h(q,p) = q_a - f(p_u)\)
are regular, since the regularity equation (\ref{eqn:regularity-matrix-acrobot})
becomes strictly positive.
A momentum-based constraint approximates human motion may still be likely to
inject energy into the acrobot.

To design the acrobot constraint, let us imagine what a human does on
a seated swing.
When they are moving forward they extend their legs, and when they are moving
backward they retract them. 
In reality, the motion is slightly more complicated than that.:
as the swing gains speed, the person leans their body back while
extending their legs.
This allows them to bring their legs higher, shortening the distance
from their center of mass to the pivot and adding more energy to the swing.
When the swing moves backward, they sit up and fully retract their legs
underneath them.

Now imagine we have affixed the torso of the acrobot to the swing's rope, so 
that the legs can pivot both forwards and backwards.  
In this case, the legs would pivot in the direction of motion
proportionally to the swing's speed.
One VNHC which emulates this process is \(q_a = \bar{q}_a\arctan_2( I p_u)\), as
in Figure \ref{fig:qa-arctan}.
Here, \(\bar{q}_a \in ]0,\frac{2 Q_a}{\pi}]\) so that 
\(q_a \in [-Q_a, Q_a]\) and \(I > 0 \in \R\) is a fixed control parameter.

This constraint is not a perfect emulation of a gymnast's movement -- during
a giant the gymnast's legs are almost completely extended, while the constraint
retracts the legs partially during rotations.
However, the motion looks similar enough that this constraint should provide a
decent foundation for injecting energy into the acrobot.
It is for this reason that we choose our acrobot's constraint to be
\begin{equation}\label{eqn:acrobot-constraint}
    h(q,p) = q_a - \bar{q}_a \arctan_2(I p_u)
    .
\end{equation}

\begin{figure}
    \centering
    \includestandalone[width=0.5\textwidth]{images/qa_arctan}
    \caption{The acrobot constraint \(q_a = \bar{q}_a \arctan_2(I p_u)\).}
    \label{fig:qa-arctan}
\end{figure}

Let us now compute the constrained dynamics for this
(\ref{eqn:acrobot-constraint}).
% TODO: should I do this for a generic h(q_a,p_u)? That's how I did it in my
% notes...
Note that \(dh_q = \begin{bmatrix}0 & 1\end{bmatrix}\tpose\), while
\[
    dh_{p_u} = \frac{-\bar{q}_a I}{1 + I^2 p_u^2}
    .
\]
Applying these to (\ref{eqn:g-qupu}), we get the solution for \(p_a\) on the
constraint manifold as
\[
    p_a(q_u,p_u) = \frac{
        (1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
    }{ml^2(3+2c_a)(1+I^2 p_u^2)}
    .
\]

\section{Proving the Acrobot Gains Energy}
% TODO: Prove this in sections
